---
title: "Quality Control in TPMS"
author: "Alisha Monsibais"
date: "2024-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Summary 
Online documentation on project:
https://docs.google.com/presentation/d/1KE5qg0QTPJgwDhP95Hefhn3OoLmeRQx9rLEfFUiybwE/edit#slide=id.p

Project: B28B was grown to log phase 0.3 and then phage SBP2@2 was added at an MOI of 10. Timepoints were collected at time 0, 10, 40, 70, 110 mins. This was ran in duplication with B28B. RNA stability reagent (Qiagen) was used to process the pellet and RNA extraction (Qiagen - Mini) was used with the DNAse I digest from Invitrogen. AccuBlue and Tape Station was ran for quality control. Samples were overnighted on dry ice to SeqCenter for lib prep and sequencing.

This script is for the bacteria analysis strain B28B using trycycler assembly and Bakta annotations



```{r}
#loading libraries 
library(DESeq2)
library(pheatmap)
library(dplyr)
library(tibble)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(ggrepel)
```


```{r loading and wrangling data}
#loading count data aligned to bacteria trycycler/bakta genome 
count_B28B <- read.delim("bacteria_counts_Bakta.txt", skip = 1,sep ="", header =TRUE)

#cleaned data with headers for geneid and sample names (column names)
count_data <-  count_B28B %>%
  mutate(Geneid = sub(".*(peg\\.\\d+)", "\\1", Geneid)) %>%
  rename_with(~ gsub("(_rep\\d+).*", "\\1", .), starts_with("B28B") | starts_with("Ph")) %>%
  select(-Start, -End, -Chr, -Strand) %>%
  column_to_rownames("Geneid")

```


Calculating TPMs
```{r}

# Step 1: Convert gene length to kilobases (for RPK calculation)
count_data  <- count_data  %>%
  mutate(length_kb = Length / 1000)  # Convert length to kilobases

# Assuming count_data has a 'length_kb' column already created
rpk_df <- count_data %>%
  mutate(across(matches("^(B28B*|Ph*)"), ~ . / length_kb, .names = "RPK_{col}"))

# Step 3: Calculate the per-sample scaling factor (sum of all RPKs in each sample)
scaling_factors <- colSums(rpk_df %>% select(starts_with("RPK")))

# Step 4: Calculate TPM for each sample
tpm_df <- rpk_df %>%
  mutate(across(starts_with("RPK"), ~ . / scaling_factors[[cur_column()]] * 1e6, .names = "TPM_{col}"))

# Optional: Clean up to keep only TPM columns and gene IDs
tpm_df <- tpm_df %>%
  select(starts_with("TPM"))

tpm_df_clean <- tpm_df %>%
  rename_with(~ gsub("^TPM_RPK_", "", .), starts_with("TPM_RPK"))

glimpse(tpm_df_clean)

```

```{r merges meta data}

#Metadata - samples with detail of samples 
meta_data <- read.csv(file = '230223_Metadata.csv', sep = ',', header = TRUE)
#remakes the Locustag as the rownames 
rownames(meta_data) <- meta_data$Sample
meta_data$Timepoint <- as.character(as.numeric(meta_data$Timepoint))
#changes the time points to character so the data is not "continuous" but "discrete"
#removed the Locustag column
head(metadata)

#match function - the row names of the metadata is the vector in the order that we want, so this will be the first argument, while the column names of the count or rpkm data is the vector to be reordered, which is the second argument 
genomic_idx <- match(rownames(meta_data), colnames(tpm_df_clean))
genomic_idx
#genomic_idx represents how to re-order the column in our counts data to be identical to the row names in metadata

# Reorder the counts data frame to have the sample names in the same order as the metadata data frame
counts_tpm_ordered  <- tpm_df_clean[ , genomic_idx]

#The rows of the metadata must exactly match the columns of the counts data, count data was reformed with match function
all(colnames(counts_tpm_ordered) == rownames(meta_data))
#Should come back TRUE

#This dataset has 4,413 (Seq Center analysis with Prokka annotation was 4,205 genes) genes, a typical free living bacteria could have a range of 1,500–7,500 genes according to the 'bacteria genome' wiki page. Thus, the RNAseq dataset makes biological sense

meta_data <- meta_data %>%
  unite(Exp, c(Timepoint, Condition), sep = "_", remove = FALSE)

```

##### 1. PCA analysis 1 
```{r}
t_counts_tpm_ordered <- counts_tpm_ordered %>% t()

zero_var_cols <- apply(t_counts_tpm_ordered, 2, function(x) sd(x, na.rm = TRUE) == 0)
sum(zero_var_cols)
t_counts_tpm_ordered_filtered <- t_counts_tpm_ordered[, !zero_var_cols]

# Run PCA
pca_result <- prcomp(t_counts_tpm_ordered_filtered, scale. = TRUE)

# Merge metadata with principal components
merged_data <- cbind(meta_data, as.data.frame(pca_result$x))


#Visualize PCA: Plot the PCA results to visualize the principal components.
# Scatter plot of the first two principal components
# Extract the percentage of variance explained

variance_explained <- round(100 * pca_result$sdev^2 / sum(pca_result$sdev^2), 2)

# Assuming Timepoint is a character variable, convert it to a factor with ordered levels
# Assuming Timepoint is a character variable
merged_data$Timepoint <- factor(merged_data$Timepoint, levels = c("0", "10", "40", "70", "110"))


# Plot with metadata information
#construct basic plot x=PC1, y= PC2
TMP_g <- ggplot(merged_data, aes(PC1, PC2)) + 
      geom_point(size=3, aes(col= Timepoint, fill = Timepoint, shape = Condition)) +
      scale_color_manual(values = c("#F8BBD0", "#F06292", "#E91E63", "#C21858", "#880E4F")) +
      geom_text_repel(aes(label=Sample), direction = "both", nudge_y=1.7, point.padding = 0.6, box.padding = 0.6, min.segment.length = unit(0.2, 'lines'), size=2.5) + 
  theme_linedraw()+
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab(paste0("PC1: ", variance_explained[1], "% Variance")) +
    ylab(paste0("PC2: ", variance_explained[2], "% Variance"))


show(TMP_g)

```

2. Heatmap

```{r heatmap}
#Step 1: Calculate Variance and Select Top 100 Genes

# Calculate variance for each gene across samples
gene_variances <- apply(tpm_df_clean, 1, var)  # Directly on row names as genes are in rownames

# Select the top 100 most variable genes
top_100_genes <- names(sort(gene_variances, decreasing = TRUE))[1:100]

# Subset the TPM dataframe to keep only the top 100 variable genes
tpm_top_100 <- tpm_df_clean[top_100_genes, ]  # Use row names directly to subset


#Step 2: Convert to Matrix for Heatmap: Now you can directly use this subset for your heatmap, as the gene IDs are already row names.
myMatrix_for_heatmap <- as.matrix(tpm_top_100)  # Convert to matrix format if needed

#Step 3: Create the Heatmap: Then, you can proceed with your heatmap code as you provided:
# Define column annotations
column_annotation <- meta_data[, c("Timepoint", "Condition"), drop = FALSE]

# Specify colors for annotations
annotation_colors <- list(
  Condition = c(Bacteria = "black", Phage = "#B7167E"), 
  Timepoint = c("0" = "gray100", "10" = "gray88", "40" = "gray77", "70" = "gray63", "110" = "gray48")
)

# Save heatmap to file
png('pheatmap_100variable_genes.png', units = "in", width = 8, height = 6, res = 600)
phm <- pheatmap(myMatrix_for_heatmap, 
                scale = "row",                 # Scale data by row (genes)
                cluster_rows = TRUE,           # Cluster rows (genes)
                cluster_cols = TRUE,           # Cluster columns (samples)
                show_rownames = FALSE,         # Hide gene names for clarity
                show_colnames = TRUE,          # Show sample names
                annotation_col = column_annotation,  # Add column annotations
                annotation_colors = annotation_colors, # Apply annotation colors
                border_color = NA
)
dev.off()


library(gridExtra)
library(gridExtra)
library(grid)

# Create text grobs for "A" and "B" at the top-left corner
label_A <- textGrob("A", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontsize = 16))

label_B <- textGrob("B", x = unit(0, "npc"), y = unit(1, "npc"),
                    just = c("left", "top"), gp = gpar(fontsize = 16))

# Combine pheatmap and ggplot with labels
final_qc_plot <- grid.arrange(
  arrangeGrob(phm$gtable, top = label_A),  # Add "A" to the top-left of pheatmap
  arrangeGrob(ggplotGrob(TMP_g), top = label_B),  # Add "B" to the top-left of ggplot
  ncol = 2  # Combine side by side
)

ggsave("Supp_Figure_3_Bakta.png", plot = final_qc_plot, width = 12, height = 5, dpi = 300)
ggsave("Supp_Figure_3_Bakta.pdf", plot = final_qc_plot, width = 12, height = 5, dpi = 300)
```

