---
title: "KB Testing from SBP2@2 Resistant Strains"
author: "Script by Alisha Monsibais"
date: '2024-01-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Project Summary 
B28B was subjected to eiether SBP2@2 at an MOI of 10 or SM as a control and starting OD600 of 0.3?. Culters were incubated on a shaker at ~200 rpm at 37C for 48 hours and bacteria was taken at the following times: 0m, 10m, 40m, 70m, 110m, 5h, 12h, 24h, 48h. Bacteria was streaked on to a petri plate and incubated for 20 hours. 

Single colonies were grown up in BHI ON then reset at a 1:20 dilution to achieve log phase bacteria at an OD600 of 0.1. Bacteria was swabed onto a petri plate and 5 antibiotic disk placed. Plates were incubated at 37C for 20 hours.  

The goal of this project is to track the bacteria physiology trade offs of phage resistance and when phage resistance occurs. 

Date:2/28/24

## 1. Exploratory Data Analysis 
```{r loading packages, message=FALSE, warning=FALSE, include=FALSE}
library(knitr) #nice table output 
library(tidyverse)
library(reshape2) #helps melt data - data wrangling 
library(ggpubr)
library(dplyr)
library(gcplyr) #AUC function 
library(ez)
#https://cran.r-project.org/web/packages/gcplyr/vignettes/gc01_gcplyr.pdf

```

```{r loading in data}
KB_repp <- read.csv("230811_Rep_prelim/230810_KB_measurements.csv")
KB_rep1 <- read.csv("240802_Rep_1/240802_KB_measurements.csv")
KB_rep2 <- read.csv("240816_Rep_2/240816_KB_measurements_wrepeats.csv")
```


```{r AUC Plate}

#plate 1
#Group by the "Well" variable and calculate AUC for each group
Results1 <- KB_rep1 %>% 
  mutate(measurement.mm = measurement * 10) %>%
  group_by(Condition, Timepoint, antibiotic) %>%
  mutate(Timepoint = as.numeric(as.character(Timepoint))) %>%
  select(Condition, Timepoint, antibiotic, measurement, measurement.mm) %>%
  mutate(Rep = 1) %>%
  filter(Timepoint != "1080")

Results2 <- KB_rep2 %>% 
  mutate(measurement.mm = measurement * 10) %>%
  group_by(Condition, Timepoint, antibiotic) %>%
  mutate(Timepoint = as.numeric(as.character(Timepoint))) %>%
   select(Condition, Timepoint, antibiotic, measurement, measurement.mm) %>%
  mutate(Rep =2) %>%
  filter(Timepoint != "1080")

Results3 <- KB_repp %>% 
  mutate(measurement.mm = measurement * 10) %>%
  group_by(Condition, Timepoint, antibiotic) %>%
  mutate(Timepoint = as.numeric(as.character(Timepoint))) %>%
  filter(Rep == "1") %>%
  select(Condition, Timepoint, antibiotic, measurement, measurement.mm) %>%
  mutate(Rep = 3) %>%
  filter(Timepoint != "720") %>%
  filter(Timepoint != "2880")


plate1 <- ggplot(data = Results1, aes(x = measurement.mm, y = antibiotic, fill = Condition)) +
  geom_col(position = position_dodge()) +  # Adjust width as needed
  scale_fill_manual(values = c("black", "#B7167E") ) +  # Add manual fill colors
  facet_grid(Timepoint~.) +
  labs(title = "Antibiotic Sensitivity - Prelim Rep", x = "ZOI (mm)", y = "Antibiotic") +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 8),
        axis.text.y = element_text(color = "black", size = 8),
        axis.title.x = element_text(color = "black", size = 8, face="bold"), 
        axis.title.y = element_text(color = "black", size = 8, face="bold"))

plate2 <- ggplot(data = Results2, aes(x = measurement.mm, y = antibiotic, fill = Condition)) +
  geom_col(position = position_dodge()) +  # Adjust width as needed
  scale_fill_manual(values = c("black", "#B7167E") ) +  # Add manual fill colors
  facet_grid(Timepoint~.) +
  labs(title = "Antibiotic Sensitivity - Rep1", x = "ZOI (mm)", y = "Antibiotic") +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 8),
        axis.text.y = element_text(color = "black", size = 8),
        axis.title.x = element_text(color = "black", size = 8, face="bold"), 
        axis.title.y = element_text(color = "black", size = 8, face="bold"))

plate3 <- ggplot(data = Results3, aes(x = measurement.mm, y = antibiotic, fill = Condition)) +
  geom_col(position = position_dodge()) +  # Adjust width as needed
  scale_fill_manual(values = c("black", "#B7167E") ) +  # Add manual fill colors
  facet_grid(Timepoint~.) +
  labs(title = "Antibiotic Sensitivity - Rep2", x = "ZOI (mm)", y = "Antibiotic") +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 8),
        axis.text.y = element_text(color = "black", size = 8),
        axis.title.x = element_text(color = "black", size = 8, face="bold"), 
        axis.title.y = element_text(color = "black", size = 8, face="bold"))


png('KB_Overview.png', units="in", width=10, height=5, res=600)
KB_overview <- ggarrange(plate1, plate2, plate3,
                      labels = c("A", "B", "C"),
                      ncol = 4, nrow = 1, 
                      common.legend = TRUE, legend = "right")
KB_overview
dev.off()

```


```{r Averages}
Alldata <- Results1 %>% union(Results2) %>% union(Results3) %>%
  group_by(Condition, Timepoint, antibiotic) %>%
  summarise(n = length(measurement.mm), Median = median(measurement.mm), Average = mean(measurement.mm), SD =   sd(measurement.mm), SE = sd(measurement.mm)/sqrt(length((measurement.mm))))

Alldata <- Alldata %>%
  filter(Timepoint %in% c("0", "10", "40", "70", "110")) %>%
  mutate(Timepoint = paste0(Timepoint, " mins \nPI")) %>%
  mutate(Timepoint = factor(Timepoint, levels = c("0 mins \nPI", "10 mins \nPI", "40 mins \nPI", "70 mins \nPI", "110 mins \nPI"))) %>%
  mutate(Condition = str_replace(Condition, "SBP22", "SBP2Φ2 Exposed")) %>%
  mutate(Condition = str_replace(Condition, "B28B", "Bacterial Control"))



Averages <- ggplot(data =Alldata, aes(x = Average, y = antibiotic, fill = Condition)) +
  geom_col(position = position_dodge()) +  # Adjust width as needed
    geom_errorbar(aes(xmin = Average - SD, xmax = Average + SD),
                position = position_dodge()) +  # Adjust width as needed
  scale_fill_manual(values = c("black", "#B7167E") ) +  # Add manual fill colors
  facet_grid(Timepoint~.) +
  labs( x = "ZOI (mm)", y = "Antibiotic") +
  xlim(0,40) +
  theme_linedraw() +
  theme(axis.text.x = element_text(color = "black", size = 10),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title.x = element_text(color = "black", size = 10, face="bold"), 
        axis.title.y = element_text(color = "black", size = 10, face="bold"), 
        strip.text.y = element_text(face = "bold", size = 10), 
         legend.position = "bottom")

ggsave("Fig6a_KB_Avgs.pdf", plot = Averages, width = 6, height = 8, dpi = 300)
ggsave("Fig6a_Kb_Avgs.png", plot = Averages, width = 6, height = 8, dpi = 300)


```
```{r Spliting up the data}
Alldata <- Results1 %>% union(Results2) %>% union(Results3)


#subsetting data 

subset_data_0_caz <- Alldata %>% filter(antibiotic == "caz", Timepoint == 0)
subset_data_0_mem <- Alldata %>% filter(antibiotic == "mem", Timepoint == 0)
subset_data_0_mh <- Alldata %>% filter(antibiotic == "mh", Timepoint == 0)
subset_data_0_lev <- Alldata %>% filter(antibiotic == "lev", Timepoint == 0)
subset_data_0_sxt <- Alldata %>% filter(antibiotic == "sxt", Timepoint == 0)

subset_data_10_caz <- Alldata %>% filter(antibiotic == "caz", Timepoint == 10)
subset_data_10_mem <- Alldata %>% filter(antibiotic == "mem", Timepoint == 10)
subset_data_10_mh <- Alldata %>% filter(antibiotic == "mh", Timepoint == 10)
subset_data_10_lev <- Alldata %>% filter(antibiotic == "lev", Timepoint == 10)
subset_data_10_sxt <- Alldata %>% filter(antibiotic == "sxt", Timepoint == 10)

subset_data_40_caz <- Alldata %>% filter(antibiotic == "caz", Timepoint == 40)
subset_data_40_mem <- Alldata %>% filter(antibiotic == "mem", Timepoint == 40)
subset_data_40_mh <- Alldata %>% filter(antibiotic == "mh", Timepoint == 40)
subset_data_40_lev <- Alldata %>% filter(antibiotic == "lev", Timepoint == 40)
subset_data_40_sxt <- Alldata %>% filter(antibiotic == "sxt", Timepoint == 40)

subset_data_70_caz <- Alldata %>% filter(antibiotic == "caz", Timepoint == 70)
subset_data_70_mem <- Alldata %>% filter(antibiotic == "mem", Timepoint == 70)
subset_data_70_mh <- Alldata %>% filter(antibiotic == "mh", Timepoint == 70)
subset_data_70_lev <- Alldata %>% filter(antibiotic == "lev", Timepoint == 70)
subset_data_70_sxt <- Alldata %>% filter(antibiotic == "sxt", Timepoint == 70)

subset_data_110_caz <- Alldata %>% filter(antibiotic == "caz", Timepoint == 110)
subset_data_110_mem <- Alldata %>% filter(antibiotic == "mem", Timepoint == 110)
subset_data_110_mh <- Alldata %>% filter(antibiotic == "mh", Timepoint == 110)
subset_data_110_lev <- Alldata %>% filter(antibiotic == "lev", Timepoint == 110)
subset_data_110_sxt <- Alldata %>% filter(antibiotic == "sxt", Timepoint == 110)

shapiro.test(subset_data_0_caz$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_0_caz)

shapiro.test(subset_data_0_mem$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_0_mem)

shapiro.test(subset_data_0_mh$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_0_mh)

shapiro.test(subset_data_0_lev$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_0_lev)

shapiro.test(subset_data_0_sxt$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_0_sxt)
#timepoint 0 is normally distributed when subsetted by antibiotics


results_0 <- Alldata %>% filter(Timepoint == "0") %>%
  group_by(antibiotic) %>%
  summarise(
    t_test_p_value = t.test(measurement.mm ~ Condition, var.equal = TRUE)$p.value,
    .groups = "drop"
  ) %>% mutate(fdr_corrected_p = p.adjust(t_test_p_value, method = "fdr"))  %>%
  mutate(t = 0)


shapiro.test(subset_data_10_caz$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_10_caz)

shapiro.test(subset_data_10_mem$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_10_mem)

shapiro.test(subset_data_10_mh$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_10_mh)

shapiro.test(subset_data_10_lev$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_10_lev)

shapiro.test(subset_data_10_sxt$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_10_sxt)
#timepoint 10 is normally distributed when subsetted by antibiotics

results_10 <- Alldata %>% filter(Timepoint == "10") %>%
  group_by(antibiotic) %>%
  summarise(
    t_test_p_value = t.test(measurement.mm ~ Condition, var.equal = TRUE)$p.value,
    .groups = "drop"
  ) %>% mutate(fdr_corrected_p = p.adjust(t_test_p_value, method = "fdr"))  %>%
  mutate(t = 10)

shapiro.test(subset_data_40_caz$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_40_caz)

shapiro.test(subset_data_40_mem$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_40_mem)
#data is not normally distributed

shapiro.test(subset_data_40_mh$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_40_mh)

shapiro.test(subset_data_40_lev$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_40_lev)

shapiro.test(subset_data_40_sxt$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_40_sxt)
#timepoint 10 is normally distributed when subsetted by antibiotics

results_40 <- Alldata %>% filter(Timepoint == "40") %>% filter(antibiotic != "mem") %>%
  group_by(antibiotic) %>%
  summarise(
    t_test_p_value = t.test(measurement.mm ~ Condition, var.equal = TRUE)$p.value,
    .groups = "drop"
  ) %>% mutate(fdr_corrected_p = p.adjust(t_test_p_value, method = "fdr"))  %>%
  mutate(t = 40)

wilcox.test(measurement.mm ~ Condition, data = subset_data_40_mem)

shapiro.test(subset_data_70_caz$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_70_caz)

shapiro.test(subset_data_70_mem$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_70_mem)

shapiro.test(subset_data_70_mh$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_70_mh)

shapiro.test(subset_data_70_lev$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_70_lev)

shapiro.test(subset_data_70_sxt$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_70_sxt)


results_70 <- Alldata %>% filter(Timepoint == "70") %>%
  group_by(antibiotic) %>%
  summarise(
    t_test_p_value = t.test(measurement.mm ~ Condition, var.equal = TRUE)$p.value,
    .groups = "drop"
  ) %>% mutate(fdr_corrected_p = p.adjust(t_test_p_value, method = "fdr"))  %>%
  mutate(t = 70)


shapiro.test(subset_data_110_caz$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_110_caz)

shapiro.test(subset_data_110_mem$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_110_mem)
#data is not normally distributed

shapiro.test(subset_data_110_mh$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_110_mh)

shapiro.test(subset_data_110_lev$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_110_lev)

shapiro.test(subset_data_110_sxt$measurement.mm)
leveneTest(measurement.mm ~ Condition, data = subset_data_110_sxt)


results_110 <- Alldata %>% filter(Timepoint == "110") %>%
  group_by(antibiotic) %>%
  summarise(
    t_test_p_value = t.test(measurement.mm ~ Condition, var.equal = TRUE)$p.value,
    .groups = "drop"
  ) %>% mutate(fdr_corrected_p = p.adjust(t_test_p_value, method = "fdr")) %>%
  mutate(t = 110)

wilcox.test(measurement.mm ~ Condition, data = subset_data_70_mem)


stats <- results_0 %>% union(results_10) %>% union(results_40) %>% union(results_70) %>% union(results_110)

write.csv(stats, "stats_results.csv", row.names = FALSE)

```















