---
title: "Synergy"
author: "Script by Alisha Monsibais, wetlab by Karen Lam and Allene Vo "
date: '2024-07-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Project Summary 
Bacteria was plated on a 96 well plate and exposed to a variation of phage concentration and antibiotic concentration. Control consisted of lysate controls (phage lysate and media), bacteria control (bacteria and media), DMSO control (bacteria, media and DMSO), blank (media, sm and saline), antibiotic controls, phage only controls    

Strain - B28B Stationary (1:20 reset and plates 24 before phage and antibiotic were added) 
Antibiotic - CAZ
Description - Four Bio Replicates
Date - 11/7/24

## 1. Exploratory Data Analysis 
```{r loading packages, message=FALSE, warning=FALSE, include=FALSE}
library(knitr) #nice table output 
library(tidyverse)
library(reshape2) #helps melt data - data wrangling 
library(ggpubr)
library(growthcurver) #growth curve package 
library(dplyr)
library(gcplyr) #AUC function 
library(car)
```

```{r loading in data}
Curve1 <- read.csv("241107_B28B_CAZ_SBP2@2_challenge_1.csv")
Curve2 <- read.csv("241107_B28B_CAZ_SBP2@2_challenge_2.csv")
Curve3 <- read.csv("241107_B28B_CAZ_SBP2@2_challenge_3.csv")
Curve4 <- read.csv("241107_B28B_CAZ_SBP2@2_challenge_4.csv")

```

```{r loading in maps}
platemap <- read.csv("Platemap_CAZ.csv")
```


```{r forloop for Data wrangling}

# Assuming you have a list of data frames like Curve1, Curve2, Curve3, etc.
curve_list <- list(Curve1, Curve2, Curve3, Curve4)

# Define the fix vector
fix <- c("00:00:00","00:10:00","00:20:00", "00:30:00", "00:40:00", "00:50:00")

# Define a function to perform the required operations
process_curve <- function(curve, platemap) {
  Time <- curve[, 1]
  Time <- Time[-c(1:6)]
  Time <- append(fix, Time, after = length(fix))
  
  # Change vector of Time from HH:MM:SS to numeric hour format
  Time <- sapply(strsplit(Time, ":"), function(x) {
    x <- as.numeric(x); x[1] + x[2]/60 + x[3]/3600
  })
  
  curve$Time <- NULL
  curve <- cbind(Time, curve)
  
  # Melt the data frame
  curve_melt <- melt(curve, id=("Time"), variable.name= "Well", value.name = "OD600")
  
  # Convert OD600 to numeric
  curve_melt$OD600 <- as.numeric(as.character(curve_melt$OD600))
  
  # Perform inner join with platemap
  annotated_curve <- inner_join(curve_melt, platemap, by='Well')
  
  return(annotated_curve)
}

# Use a for loop to process each data frame in the list
for (i in seq_along(curve_list)) {
  curve_list[[i]] <- process_curve(curve_list[[i]], platemap)
}

# Assuming your list is named 'your_list' - this is in the order of the curves 1-4
names(curve_list) <- c("Replicate 1", "Replicate 2", "Replicate 3", "Replicate 4")


```

```{r plots the overview of the 96 well plate}
# Creating a list to store ggplots
ggplot_list <- list()

# Iterating over the list
for (i in names(curve_list)) {
  df <- curve_list[[i]]
  plot_title <- paste("Synergy of SBP2Φ2 and CAZ -", i)
  ggplot_list[[i]] <- ggplot(df, aes(x = Time, y = OD600, color = Condition)) +
    ggtitle(plot_title) +
    geom_line() +
    labs(x = "Time (Hours)", y = "OD600nm") +
    theme(
      plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
      axis.title.x = element_text(size = 9, face = "bold"),
      axis.title.y = element_text(size = 9, face = "bold"), 
      axis.text.x = element_text(size = 5), 
      axis.text.y = element_text(size = 5)) +
    facet_grid(Letter ~ Column) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
  # Save the plot as a PNG file
  ggsave(paste0("figures/B28B_CAZ_plot_241107_", i, ".png"), ggplot_list[[i]], width = 7, height = 4, units = "in")
}

# Create the ggarrange object using ggplot_list
Plate_96 <- ggarrange(plotlist = ggplot_list,
                      labels = c("A", "B", "C", "D"),
                      ncol = 2, nrow = 2, 
                      common.legend = TRUE, legend = "right")

# Save the plot as a PNG file
ggsave("figures/B28B_CAZ_PlateOverview_241107.png", Plate_96, width = 12, height = 6, units = "in", dpi = 300)

```


```{r Synergy Plate}
library(gcplyr)

#plate 1
#Group by the "Well" variable and calculate AUC for each group
AUC_results_1 <- curve_list[["Replicate 1"]] %>% filter(Time <= 30) %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  summarise(uncorrected_AUC = auc(y=OD600, x=Time)) %>%
  ungroup() %>%
  inner_join(platemap, by='Well') %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(Blank = mean(uncorrected_AUC[Condition == 'Blank'])) %>%
  mutate(AUC = uncorrected_AUC - Blank) %>% 
  mutate(Bacteria_AUC = AUC[Condition == 'Bacterial_Control']) %>%
  mutate(growthpercent = (1-(Bacteria_AUC - AUC)/Bacteria_AUC)*100) %>%
  as.data.frame() %>% 
  mutate(CAZ = factor(CAZ, levels = c("0","0.125", "0.25", "0.5", "1", "2", "4", "8", "16", "32"))) %>%
  mutate(Phage = factor(Phage, levels = c("1.0E+08", "1.0E+07", "1.0E+06", "1.0E+05", "1.0E+04", "1.0E+03", "0")))

plate1 <- ggplot(data = AUC_results_1) + 
  geom_tile(aes(x = Phage, y = CAZ, fill = growthpercent), color = "black") +  # Adjust width and height as needed +
  scale_fill_gradient(low = "white", high = "#B7167E") +  # Adjust the color gradient as needed
  labs(title = "Synergy of SBP2Φ2 and CAZ against B28B", 
       subtitle = "Growth Percentage over 30 hours \n Rep 1", 
       x = "SBP2Φ2 (PFU/mL)", y = "CAZ (ug/mL)") +
  theme_minimal() + # Adjust the theme as needed
  scale_x_discrete(lim=rev) + 
  theme(plot.title = element_text(hjust = 0.5),  
  plot.subtitle = element_text(hjust = 0.5), 
  axis.text.x = element_text(angle = 45, hjust = 1)) +
   labs(fill = "Growth %")

plate1

#plate2
#Group by the "Well" variable and calculate AUC for each group
AUC_results_2 <- curve_list[["Replicate 2"]] %>% filter(Time <= 30) %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  summarise(uncorrected_AUC = auc(y=OD600, x=as.numeric(Time))) %>%
  ungroup() %>%
  inner_join(platemap, by='Well') %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(Blank = mean(uncorrected_AUC[Condition == 'Blank'])) %>%
  mutate(AUC = uncorrected_AUC - Blank) %>% 
  mutate(Bacteria_AUC = AUC[Condition == 'Bacterial_Control']) %>%
  mutate(growthpercent = (1-(Bacteria_AUC - AUC)/Bacteria_AUC)*100) %>%
  as.data.frame() %>% 
  mutate(CAZ = factor(CAZ, levels = c("0","0.125", "0.25", "0.5", "1", "2", "4", "8", "16", "32"))) %>%
  mutate(Phage = factor(Phage, levels = c("1.0E+08", "1.0E+07", "1.0E+06", "1.0E+05", "1.0E+04", "1.0E+03", "0")))


plate2 <- ggplot(data = AUC_results_2) + 
  geom_tile(aes(x = Phage, y = CAZ, fill = growthpercent), color = "black") +  # Adjust width and height as needed +
  scale_fill_gradient(low = "white", high = "#B7167E") +  # Adjust the color gradient as needed
  labs(title = "Synergy of SBP2Φ2 and CAZ against B28B", 
       subtitle = "Growth Percentage over 30 hours \n Rep 2", 
       x = "SBP2Φ2 (PFU/mL)", y = "CAZ (ug/mL)") +
  theme_minimal() + # Adjust the theme as needed
  scale_x_discrete(lim=rev) + 
  theme(plot.title = element_text(hjust = 0.5),  
  plot.subtitle = element_text(hjust = 0.5), 
  axis.text.x = element_text(angle = 45, hjust = 1)) +
   labs(fill = "Growth %")

plate2

#Plate 3
#Group by the "Well" variable and calculate AUC for each group
AUC_results_3 <- curve_list[["Replicate 3"]] %>% filter(Time <= 30) %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  summarise(uncorrected_AUC = auc(y=OD600, x=as.numeric(Time))) %>%
  ungroup() %>%
  inner_join(platemap, by='Well') %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(Blank = mean(uncorrected_AUC[Condition == 'Blank'])) %>%
  mutate(AUC = uncorrected_AUC - Blank) %>% 
  mutate(Bacteria_AUC = AUC[Condition == 'Bacterial_Control']) %>%
  mutate(growthpercent = (1-(Bacteria_AUC - AUC)/Bacteria_AUC)*100) %>%
  as.data.frame() %>% 
  mutate(CAZ = factor(CAZ, levels = c("0","0.125", "0.25", "0.5", "1", "2", "4", "8", "16", "32"))) %>%
  mutate(Phage = factor(Phage, levels = c("1.0E+08", "1.0E+07", "1.0E+06", "1.0E+05", "1.0E+04", "1.0E+03", "0")))

plate3 <- ggplot(data = AUC_results_3) + 
  geom_tile(aes(x = Phage, y = CAZ, fill = growthpercent), color = "black") +  # Adjust width and height as needed +
  scale_fill_gradient(low = "white", high = "#B7167E") +  # Adjust the color gradient as needed
  labs(title = "Synergy of SBP2Φ2 and CAZ against B28B", 
       subtitle = "Growth Percentage over 30 hours \n Rep 3", 
       x = "SBP2Φ2 (PFU/mL)", y = "CAZ (ug/mL)") +
  theme_minimal() + # Adjust the theme as needed
  scale_x_discrete(lim=rev) + 
  theme(plot.title = element_text(hjust = 0.5),  
  plot.subtitle = element_text(hjust = 0.5), 
  axis.text.x = element_text(angle = 45, hjust = 1)) +
   labs(fill = "Growth %")

plate3

#Plate 4
#Group by the "Well" variable and calculate AUC for each group
AUC_results_4 <- curve_list[["Replicate 4"]] %>% filter(Time <= 30) %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  summarise(uncorrected_AUC = auc(y=OD600, x=as.numeric(Time))) %>%
  ungroup() %>%
  inner_join(platemap, by='Well') %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(Blank = mean(uncorrected_AUC[Condition == 'Blank'])) %>%
  mutate(AUC = uncorrected_AUC - Blank) %>% 
  mutate(Bacteria_AUC = AUC[Condition == 'Bacterial_Control']) %>%
  mutate(growthpercent = (1-(Bacteria_AUC - AUC)/Bacteria_AUC)*100) %>%
  as.data.frame() %>% 
  mutate(CAZ = factor(CAZ, levels = c("0","0.125", "0.25", "0.5", "1", "2", "4", "8", "16", "32"))) %>%
  mutate(Phage = factor(Phage, levels = c("1.0E+08", "1.0E+07", "1.0E+06", "1.0E+05", "1.0E+04", "1.0E+03", "0")))

plate4 <- ggplot(data = AUC_results_4) + 
  geom_tile(aes(x = Phage, y = CAZ, fill = growthpercent), color = "black") +  # Adjust width and height as needed +
  scale_fill_gradient(low = "white", high = "#B7167E") +  # Adjust the color gradient as needed
  labs(title = "Synergy of SBP2Φ2 and CAZ against B28B", 
       subtitle = "Growth Percentage over 30 hours \n Rep 3", 
       x = "SBP2Φ2 (PFU/mL)", y = "CAZ (ug/mL)") +
  theme_minimal() + # Adjust the theme as needed
  scale_x_discrete(lim=rev) + 
  theme(plot.title = element_text(hjust = 0.5),  
  plot.subtitle = element_text(hjust = 0.5), 
  axis.text.x = element_text(angle = 45, hjust = 1)) +
   labs(fill = "Growth %")

plate4


png('figures/B28B_CAZ_AUC_Overview_241107_30hours.png', units="in", width=18, height=6, res=600)
Plate_96 <- ggarrange(plate1, plate2, plate3, plate4,
                      labels = c("A", "B", "C", "D"),
                      ncol = 4, nrow = 1, 
                      common.legend = TRUE, legend = "right")
Plate_96
dev.off()

```


```{r}

Plate1_96 <- curve_list[["Replicate 1"]] %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(CAZ = factor(CAZ, levels = c("32", "16", "8", "4", "2", "1", "0.5", "0.25", "0.125","0"))) %>%
  mutate(Phage = factor(Phage, levels = c("0", "1.0E+03", "1.0E+04", "1.0E+05", "1.0E+06", "1.0E+07", "1.0E+08")))

plate_wells_1 <- ggplot(Plate1_96, aes(x=Time, y=OD600, color = Condition))+
  ggtitle("B28B: CAZ and SBP2Φ2 Synergy", 
          subtitle = "Rep 1") +
  geom_line() +
  labs(x="Time(Hours)", y ="OD600nm") +
  #scale_x_discrete(lim=rev) + 
  theme(
    plot.title = element_text(size=12, hjust = 0.5, face="bold"),
    axis.title.x = element_text(size=9, face="bold"),
    axis.title.y = element_text(size=9, face="bold"), 
    axis.text.x = element_text(size=5), 
    axis.text.y = element_text(size=5))+
  facet_grid(CAZ~Phage) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
plate_wells_1
  
png('figures/B28B_CAZ_plate1_241107.png', units="in", width=13, height=6, res=600)  
Plate_1 <- ggarrange(plate_wells_1, plate1, labels = c("A", "B"))
Plate_1
dev.off()

#plate2
Plate2_96 <- curve_list[["Replicate 2"]] %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(CAZ = factor(CAZ, levels = c("32", "16", "8", "4", "2", "1", "0.5", "0.25", "0.125","0"))) %>%
  mutate(Phage = factor(Phage, levels = c("0", "1.0E+03", "1.0E+04", "1.0E+05", "1.0E+06", "1.0E+07", "1.0E+08")))


plate_wells_2 <- ggplot(Plate2_96, aes(x=Time, y=OD600, color = Condition))+
  ggtitle("B28B: CAZ and SBP2Φ2 Synergy", 
          subtitle = "Rep 2") +
  geom_line() +
  labs(x="Time(Hours)", y ="OD600nm") +
  #scale_x_discrete(lim=rev) + 
  theme(
    plot.title = element_text(size=12, hjust = 0.5, face="bold"),
    axis.title.x = element_text(size=9, face="bold"),
    axis.title.y = element_text(size=9, face="bold"), 
    axis.text.x = element_text(size=5), 
    axis.text.y = element_text(size=5))+
  facet_grid(CAZ~Phage) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
plate_wells_2
  
png('figures/B28B_CAZ_plate2_241107.png', units="in", width=13, height=6, res=600)  
Plate_2 <- ggarrange(plate_wells_2, plate2, labels = c("A", "B"))
Plate_2
dev.off()

#plate3
Plate3_96 <- curve_list[[3]] %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(CAZ = factor(CAZ, levels = c("32", "16", "8", "4", "2", "1", "0.5", "0.25", "0.125","0"))) %>%
  mutate(Phage = factor(Phage, levels = c("0", "1.0E+03", "1.0E+04", "1.0E+05", "1.0E+06", "1.0E+07", "1.0E+08")))


plate_wells_3 <- ggplot(Plate3_96, aes(x=Time, y=OD600, color = Condition))+
  ggtitle("B28B: CAZ and SBP2Φ2 Synergy", 
          subtitle = "Rep 3") +
  geom_line() +
  labs(x="Time(Hours)", y ="OD600nm") +
  #scale_x_discrete(lim=rev) + 
  theme(
    plot.title = element_text(size=12, hjust = 0.5, face="bold"),
    axis.title.x = element_text(size=9, face="bold"),
    axis.title.y = element_text(size=9, face="bold"), 
    axis.text.x = element_text(size=5), 
    axis.text.y = element_text(size=5))+
  facet_grid(CAZ~Phage) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
plate_wells_3

png('figures/B28B_CAZ_plate3_241107.png', units="in", width=13, height=6, res=600)  
Plate_3 <- ggarrange(plate_wells_3, plate3, labels = c("A", "B"))
Plate_3
dev.off()

#plate4
Plate4_96 <- curve_list[[4]]  %>% 
  group_by(Well) %>%
  mutate(Time = round(Time, 1)) %>%
  filter(!Condition %in% c("UPW", "DMSO_Control", "Lysate_Control")) %>%
  mutate(CAZ = factor(CAZ, levels = c("32", "16", "8", "4", "2", "1", "0.5", "0.25", "0.125","0"))) %>%
  mutate(Phage = factor(Phage, levels = c("0", "1.0E+03", "1.0E+04", "1.0E+05", "1.0E+06", "1.0E+07", "1.0E+08")))


plate_wells_4 <- ggplot(Plate4_96, aes(x=Time, y=OD600, color = Condition))+
  ggtitle("B28B: CAZ and SBP2Φ2 Synergy", 
          subtitle = "Rep 4") +
  geom_line() +
  labs(x="Time(Hours)", y ="OD600nm") +
  #scale_x_discrete(lim=rev) + 
  theme(
    plot.title = element_text(size=12, hjust = 0.5, face="bold"),
    axis.title.x = element_text(size=9, face="bold"),
    axis.title.y = element_text(size=9, face="bold"), 
    axis.text.x = element_text(size=5), 
    axis.text.y = element_text(size=5))+
  facet_grid(CAZ~Phage) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
plate_wells_4
  
  
png('figures/B28B_CAZ_plate4_241107.png', units="in", width=13, height=6, res=600)  
Plate_4 <- ggarrange(plate_wells_4, plate4, labels = c("A", "B"))
Plate_4
dev.off()

```

```{r}
rep_1 <- AUC_results_1 %>%
  select(CAZ, Phage, Bacteria, AUC, Condition) %>%
  mutate(rep = 1)

#excluded two due to antibiotic control 

rep_2 <- AUC_results_3 %>%
  select(CAZ, Phage, Bacteria, AUC, Condition) %>%
  mutate(rep = 2)

rep_3 <- AUC_results_4 %>%
  select(CAZ, Phage, Bacteria, AUC, Condition) %>%
  mutate(rep = 3)


#averages the data 
combined_data <- reduce(list(rep_1, rep_2, rep_3), union) %>%
  filter(!Condition == "Blank") %>%
  group_by(CAZ, Phage, Bacteria, Condition) %>%
  summarize(
    Avg_AUC = mean(AUC, na.rm = TRUE),
    SD_AUC = sd(AUC, na.rm = TRUE), 
    n = length(AUC)
  ) %>%
  mutate(
    Avg_AUC = signif(Avg_AUC, digits = 3),
    SD_AUC = signif(SD_AUC, digits = 3)
  ) %>%  ungroup() %>%
  mutate(Bacteria_AUC = Avg_AUC[Condition == 'Bacterial_Control']) %>%
  mutate(growthpercent = (1-(Bacteria_AUC - Avg_AUC)/Bacteria_AUC)*100)

Avg_plate <- ggplot(data = combined_data) + 
  geom_tile(aes(x = Phage, y = CAZ, fill = growthpercent), color = "black") +  # Adjust width and height as needed +
  scale_fill_gradient(
  low = "white", 
  high = "#004900", 
  limits = c(0, 107),  # Assuming growth percentage ranges from 0 to 100
  breaks = seq(0, 100, by = 25),  # Define legend breaks if necessary
  labels = function(x) paste0(x, "%")  # Append '%' to labels
)+
  labs(x = "SBP2Φ2 (PFU/mL)", y = "CAZ (ug/mL)") +
  theme_minimal() + # Adjust the theme as needed
  scale_x_discrete(lim=rev) + 
  theme(plot.title = element_text(hjust = 0.5),  
  plot.subtitle = element_text(hjust = 0.5), 
  axis.text.x = element_text(angle = 45, hjust = 1), 
  legend.position = "bottom") +
   labs(fill = "Growth %")

Avg_plate

png('figures/Figure7B_StationarySynergy_30hours_bottomlegend.png', units="in", width=5, height=6, res=600)
Avg_plate
dev.off()

pdf('figures/Figure7B_StationarySynergy_30hours_bottomlegend.pdf', width=5, height=6)
Avg_plate
dev.off()


```
```{r Growth Curve Analysis}

Rep1_GC <- Plate1_96 %>%
  mutate(rep = 1)
Rep2_GC <- Plate3_96 %>%
  mutate(rep = 2)
Rep3_GC <- Plate4_96 %>%
  mutate(rep = 3)

#excluded two due to antibiotic control 

# Summarize first
RawData <- Rep1_GC %>%
  union(Rep2_GC) %>%
  union(Rep3_GC) %>% 
  filter(Time <= 30) %>% 
  filter(CAZ %in% c(0, 8)) %>%
  filter(as.character(Phage) %in% c("0", "1.0E+08"))

#Extract blank information 
blank_data <- RawData %>%
  filter(Condition == "Blank") %>%
  group_by(Time) %>%
  summarize(BlankAvg = mean(OD600, na.rm = TRUE)) %>%
  ungroup()

# Subtract blank from raw OD600 readings
GrowthCurveData <- RawData %>%
  left_join(blank_data, by = "Time") %>%
  mutate(CorrOD600 = OD600 - BlankAvg) %>%
  select(-BlankAvg) %>% filter(Condition != "Blank")

# Recreate Condition now
GrowthCurveData <- GrowthCurveData %>% 
  group_by(Time, CAZ, Phage, Bacteria) %>%
  summarize(
    Avg = mean(CorrOD600, na.rm = TRUE),
    SD = sd(CorrOD600, na.rm = TRUE),
    n = n(), .groups = "drop"
  ) %>%
  mutate(Condition = case_when(
    (CAZ == 8 & Phage == "0") ~ "Antibiotic Only",
    (CAZ == 8 & Phage == "1.0E+08") ~ "Combination",
    (CAZ == 0 & Phage == "0" & Bacteria == "Bacteria") ~ "Bacterial Control",
    (CAZ == 0 & Phage == "1.0E+08") ~ "Phage Only"
  ))  %>%
   mutate(Condition = factor(Condition, levels = c(
    "Bacterial Control",   # 1st in legend
    "Antibiotic Only",     # 2nd in legend
    "Phage Only",          # 3rd in legend
    "Combination"          # 4th in legend
  )))


GrowthCurve <- ggplot(GrowthCurveData, aes(x = Time, y = Avg, color = Condition)) +
  geom_ribbon(aes(ymin = Avg - SD, ymax = Avg + SD), fill = "#E3E3E3", linetype = 0) +
  geom_line(size = 1.5) +
  scale_color_manual(values = c(
    "Bacterial Control" = "black",
    "Antibiotic Only" = "darkorange2",
    "Phage Only" = "#B7167E",
    "Combination" = "purple"
  )) +
  theme_classic2() +
  ylim(0,1.6)+
  labs(
    x = "Time (Hours)",
    y = "OD600 (nm)"
  ) +
  theme(
    legend.background = element_rect(fill = "white", color = "black"),
    legend.position = "bottom"
  )
GrowthCurve
png('figures/Figure5D_StationarySynergy_GrowthCurves.png', units="in", width=8.5, height=6, res=600)
GrowthCurve
dev.off()

pdf('figures/Figure5D_StationarySynergy_GrowthCurves.pdf', width=8.5, height=6)
GrowthCurve
dev.off()

```



```{r Growth Curve Statistics}
#Raw data is already filtered for the growth curves in the above chunck
GC_AUC <- RawData %>% group_by(Well, rep) %>%
  summarise(uncorrected_AUC = auc(y=OD600, x=as.numeric(Time))) %>%
  ungroup() %>%
  inner_join(platemap, by='Well') %>%
  mutate(Blank = mean(uncorrected_AUC[Condition == 'Blank'])) %>%
  mutate(AUC = uncorrected_AUC - Blank) %>% 
  filter(!Condition %in% c("Blank"))

#establishing the model for ANOVA check
model <- aov(AUC ~ Condition, data = GC_AUC)

#check ANOVA assumptions!
#1. Checking normality of residual
# Extract residuals
residuals_model <- residuals(model)

# Plot a Q-Q plot
qqnorm(residuals_model)
qqline(residuals_model, col = "red")
#points follow a straight line! thus, normal 

#Checking Homogeneity of Variances
# Levene's test (less sensitive to departures from normality)
leveneTest(AUC ~ Condition, data = GC_AUC)

# If you prefer Bartlett's test:
bartlett.test(AUC ~ Condition, data = GC_AUC)
#p-value 0.221 > 0.05
#data do not violate the assumption of homogeneity of variances

plot(fitted(model), residuals_model)
abline(h=0, col="red")
#residuals should appear as a random scatter around zero


#running the ANOVA
result <- aov(AUC ~ Condition, data = GC_AUC)
summary(result)

sink("ANOVA_output.txt")
summary(result)
sink()

#Running Post-Hoc Test
sum_data <- TukeyHSD(result)
#annotation are going to be added manually in adobe


sink("TukeyHSD_output.txt")
print(sum_data)
sink()

```




```{rextra formating for GC}
axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    axis.text.x = element_text(size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),
    legend.background = element_rect(fill = "white", color = "black"),
    legend.box.background = element_rect(fill = "gray90"),
    legend.text = element_text(size = 12),      # Adjusts legend text size
    legend.title = element_text(size = 12),
    legend.position = "right"
```